DeepPicar:​ ​A​ ​Low-cost​ ​Deep​ ​Neural​ ​Network-based​ ​Autonomous​ ​Car

### 2017-10-07 paper/evaluation.tex
Changes to evaluation

The Raspberry Pi 3 Model B platform used in DeepPicar equips a Boardcom
BCM2837 SoC, which has a quad-core 64bit ARM Cortex-A53 cluster,
running at up to 1.2GHz. Each core has 16K private I\&D caches, and all
cores share a 512KB L2 cache.
The chip also includes Broadcom's Videocore IV
GPU, although we did not use the GPU in our evaluation due to the lack
of sofware support (Tensorflow is not compatible with the Raspberry Pi's GPU).
For software, we use Ubuntu MATE 16.04, Tensorflow 1.1 and python
2.7. We disabled DVFS (dynamic voltage frequency scaling) and
configure the clock speed of each core at the maximum 1.2GHz.

@@

First, as expected, we find that the inference operation
dominates the control loop execution time, accounting about 85\% of
the execution time.
Second, more importantly, we also find that the measured average
execution time of a single control loop is 42.67 ms, or 23.4 Hz and
the 99 percentail time is 50.70ms.
This means that the DeepPicar can operate
at about 20 Hz control frequency in real-time using only the on-board
Raspberry Pi 3 computing platform, without needing any remote computing
resources. We consider these results respectable given the complexity
of the deep neural network  and the fact that the inference operation
performed by Tensorflow only utilizes the CPU cores of the
Raspberry Pi 3 (its GPU is not supported by Tensorflow.)

@@

Figure~\ref{fig:perf-vs-corecnt} shows the average execution time of
the control loop as we vary the number of cores used by
Tensorflow. As expected as we assign more cores, the average execution
time decreases---from 61.96 ms on a single core to 42.67 ms on four
cores (a 30\% improvement). However, the improvement is far from ideal
linear scaling. In particular, from 2 cores to 3 cores, the
improvement is mere 2.38ms (or 4\%). In short, we find that the
scalability of DeepPicar's deep neural network is not ideal on the
platform. We do not know whether it is due to the limitations of
Tensorflow's multicore implementation or it's the model's inherent
characteristics.

@@

In this experiment, we launch multiple instances of DeepPicar's DNN
model at the same time and measure its impact to their inference
timings. In particular, we are interested in how shared resource
contention affects inference timing.

@@

Figure~\ref{fig:perf-vs-modelcnt} shows the results. In the figure, the
X-axis shows the system configuration: \#of DNN models x \#of CPU
cores\/DNN. For example, '4x1' means running four DNN models each of
which is assigned to run on one CPU core, whereas '2x2' means running
two DNN models, each of which is assigned to run on two CPU
cores. The Y-axis shows the average inference timing.
Two bars in the left show the impact of co-scheduling four DNN
models. Compared to executing a single DNN model on one CPU core
(1x1), when four DNN models are co-scheduled (4x1), each model
suffers an average inference time increase of approximately 15 ms,
$\sim$24\%. On the other hand, when two DNN models, each using two CPU
cores, are co-scheduled, the average inference timing is increased by
about 7 ms, or 10\%, compared to the baseline of running one model
using two CPU cores.

@@

These increases in inference times in the co-scheduled scenarios are
expected and likely caused by contention in the shared hardware
resources such as the shared L2 cache and/or the DRAM controller.
To further analyze the source of contention, we use hardwawre
performance counters of the processor. Specifically, we measure L2
miss rates of the DNN models first in isolation and then after
co-scheduling another models. If the shared L2 cache is the primary
source of inteference, then the measured L2 miss rates would
increase. Figure~\ref{fig:l2missrate-vs-modelcnt} shows the results.
As can be see in the figure, L2 miss rates are largely unchanged
regardless whether multiple models are co-scheduled or not. This
suggests that the shared L2 cache is not the bottleneck that caused
execution time increases. In other words, DNN models appear not to be
sensitive to the shared L2 cache space. % L2 partitioning is probably
                                % not going to be useful.
Instead, we hypothesize that it is likely caused by the memory
controller---the only other major shared hardware source---where
memory requests from different CPU cores contend, which would result
in increased memory access latency. While some Intel processors
provide oncore hardware counters that can measure average memory
access latency~\cite{ye2016maracas}, we were not able to identify
whether such hardware counters exist in the BCM2837 processor of
Raspberry Pi 3 due to the lack of documentation. Instead, in the next
experiment, we use memory intensive synthetic benchmarks to test the
hypothesis.

@@

Figure~\ref{fig:} shows the normalized execution time and L2 miss-rate
of the DNN model running on the Core 0 as a function of the number of
co-scheduled memory intensive synthetic benchmarks. First, as we
increase the number of co-runners, the DNN model's execution times are
increased---by up to 9.6X---even though the DNN model is running on a
dedicated core (Core0). On the other hand, the DNN model's L2
cache-miss rates did not increase as much. This suggests that
the DNN model's exeuction increase cannot be fully explained by
increased L2 cache-misses. Instead, as we hypothesized in the previous
experiment, the increased memory pressure from the co-scheduled memory
intensive benchmarks is the primary cause of the DNN model's execution
time increase. Therefore, we conclude that DeepPicar's DNN model is
more senstive to DRAM access latency than L2 cache space.

@@

Lastly, we find that consolidating memory (DRAM) performance
intensive applications could jeopadize DNN performance because DNN
performance appears very sensitive to memory performance; we observe up
to 9.6X slowdown in DNN performance by co-scheduling synthetic memory 
bandwidth intensive applications on idle cores.

>>>>

The Raspberry Pi 3 Model B platform used in DeepPicar equips a Boardcom
BCM2837 SoC, which has a quad-core 64bit ARM Cortex-A53 cluster,
running at up to 1.2GHz. Each core has 16K private I\&D caches, and all
cores share a 512KB L2 cache.
The chip also includes Broadcom's Videocore IV
GPU, although we did not use the GPU in our evaluation due to the lack
of sofware support (Tensorflow is not compatible with the Raspberry Pi's GPU).
For software, we use Ubuntu MATE 16.04, Tensorflow 1.1 and Python
2.7. We disabled DVFS (dynamic voltage frequency scaling) and
configured the clock speed of each core at the maximum 1.2GHz.

@@

First, as expected, we find that the inference operation
dominates the control loop execution time, accounting for about 85\% of
the execution time.
Second, more importantly, we also find that the measured average
execution time of a single control loop is 42.67 ms, or 23.4 Hz and
the 99 percentail time is 50.70ms.
This means that the DeepPicar can operate
at about 20 Hz control frequency in real-time using only the on-board
Raspberry Pi 3 computing platform, without needing any remote computing
resources. We consider these results respectable given the complexity
of the deep neural network, and the fact that the inference operation
performed by Tensorflow only utilizes the CPU cores of the
Raspberry Pi 3 (its GPU is not supported by Tensorflow.)

@@

Figure~\ref{fig:perf-vs-corecnt} shows the average execution time of
the control loop as we vary the number of cores used by
Tensorflow. As expected, as we assign more cores, the average execution
time decreases---from 61.96 ms on a single core to 42.67 ms on four
cores (a 30\% improvement). However, the improvement is far from ideal
linear scaling. In particular, from 2 cores to 3 cores, the
improvement is mere 2.38ms (or 4\%). In short, we find that the
scalability of DeepPicar's deep neural network is not ideal on the
platform. We do not know whether it is due to the limitations of
Tensorflow's multicore implementation or it's the model's inherent
characteristics.

@@

In this experiment, we launch multiple instances of DeepPicar's DNN
model at the same time and measure its impact on their inference
timings. In particular, we are interested in how shared resource
contention affects inference timing.

@@

Figure~\ref{fig:perf-vs-modelcnt} shows the results. In the figure, the
X-axis shows the system configuration: \#of DNN models x \#of CPU
cores/DNN. For example, '4x1' means running four DNN models each of
which is assigned to run on one CPU core, whereas '2x2' means running
two DNN models, each of which is assigned to run on two CPU
cores. The Y-axis shows the average inference timing.
Two bars in the left show the impact of co-scheduling four DNN
models. Compared to executing a single DNN model on one CPU core
(1x1), when four DNN models are co-scheduled (4x1), each model
suffers an average inference time increase of approximately 15 ms,
$\sim$24\%. On the other hand, when two DNN models, each using two CPU
cores, are co-scheduled, the average inference timing is increased by
about 7 ms, or 10\%, compared to the baseline of running one model
using two CPU cores.

@@

These increases in inference times in the co-scheduled scenarios are
expected and likely caused by contention in the shared hardware
resources such as the shared L2 cache and/or the DRAM controller.
To further analyze the source of contention, we use hardwawre
performance counters of the processor. Specifically, we measure L2
miss rates of the DNN models first in isolation and then after
co-scheduling other models. If the shared L2 cache is the primary
source of inteference, then the measured L2 miss rates would
increase. Figure~\ref{fig:l2missrate-vs-modelcnt} shows the results.
As can be see in the figure, L2 miss rates are largely unchanged
regardless whether multiple models are co-scheduled or not. This
suggests that the shared L2 cache is not the bottleneck that caused
execution time increases. In other words, DNN models don't appear to be
sensitive to the shared L2 cache space. % L2 partitioning is probably
                                % not going to be useful.
Instead, we hypothesize that it is likely caused by the memory
controller---the only other major shared hardware source---where
memory requests from different CPU cores contend, which would result
in increased memory access latency. While some Intel processors
provide oncore hardware counters that can measure average memory
access latency~\cite{ye2016maracas}, we were not able to identify
whether such hardware counters exist in the BCM2837 processor of
Raspberry Pi 3 due to the lack of documentation. Instead, in the next
experiment, we use memory intensive synthetic benchmarks to test the
hypothesis.

@@

Figure~\ref{fig:} shows the normalized execution time and L2 miss-rate
of the DNN model running on the Core 0 as a function of the number of
co-scheduled memory intensive synthetic benchmarks. First, as we
increase the number of co-runners, the DNN model's execution times are
increased---by up to 9.6X---even though the DNN model is running on a
dedicated core (Core 0). On the other hand, the DNN model's L2
cache-miss rates did not increase as much. This suggests that
the DNN model's exeuction increase cannot be fully explained by
increased L2 cache-misses. Instead, as we hypothesized in the previous
experiment, the increased memory pressure from the co-scheduled memory
intensive benchmarks is the primary cause of the DNN model's execution
time increase. Therefore, we conclude that DeepPicar's DNN model is
more senstive to DRAM access latency than L2 cache space.

@@

Lastly, we find that consolidating memory (DRAM) performance
intensive applications could jeopadize DNN performance, because DNN
performance appears to be very sensitive to memory performance; we observe up
to 9.6X slowdown in DNN performance by co-scheduling synthetic memory 
bandwidth intensive applications on idle cores.
